---
output: github_document
---

# gie

<!-- badges: start -->
[![Lifecycle: experimental](https://lifecycle.r-lib.org/articles/figures/lifecycle-experimental.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/yannikbuhl/gie/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/yannikbuhl/gie/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

A R wrapper for Gas Infrastructure Europe's (GIE) AGSI+ and ALSI+ transparency platform API

## Collecting data on natural gas storages
[Gas Infrastructure Europe (GIE)](https://www.gie.eu/) is an organisation that represents the vast majority of operators of natural gas storage units in Europe and some non-EU countries (such as the United Kingdom and Ukraine). They provide a REST API for their AGSI+ transparency platform to retrieve data on country level, operator level and facility level. Data include information on, e.g., the filling level, overall capacity, inflow or outflow. In order to use `{gie}`, you have to register for an API key on the [AGSI+ website](https://agsi.gie.eu/).

In spring of 2022, Gas Infrastructure Europe (GIE) provided a complete makeover of its API, introducing, among changes to parameters and endpoints, the use of pagination. This package provides easy-to-use functions to download small amounts of data or use pagination to download large amounts of (historical) data. Also, it includes a function to get metadata on operators and facilities across Europe and within each country.

Get the package from CRAN:

```{r eval=FALSE}
install.packages("gie")
```

Install the development version via GitHub:

```{r eval=FALSE}
remotes::install_github("yannikbuhl/gie")
```

## Usage

The function `gie_listing()` returns a nested list with all gas facilities available, including the EIC codes for companies and facilities (to do so, do not set the 'region' and 'country' parameter and 'facilities' to `FALSE`). Also, it is possible to get parsed results (as `data.frame`) for the region and country level. Setting the 'facilities' parameter to `TRUE`, the function returns all the available facilities within a country (including company and facility EIC codes for further use in the download functions `gie_load()` and `gie_batchload()`). 

```{r eval=FALSE}

german_facilities <- gie::gie_listing(region = "Europe",
                                      country = "Germany",
                                      facilities = TRUE,
                                      apikey = apikey)

```

One main function for data download is called `gie_load()` and can be used to download data from AGSI+, either for one country, for one operator in a certain country or for a certain facility (see documentation for info on all parameters).

```{r eval=FALSE}

gasdata <- gie::gie_load(country = "de",
                         from = "2022-01-01",
                         to = "2022-03-31",
                         size = 50,
                         verbose = TRUE,
                         apikey = apikey)

```


The second download function `gie_batchload()` lets you download data for multiple countries, companies or facilities at once. Note, however, that as of yet, it is only possible to download a set of various companies for within one country; also, you can only download multiple facilities if the country and the company are fixed (due to the way the API is built).

```{r eval=FALSE} 

companydata <- gie::gie_batchload(countries = c("DE", "IE", "NL"), 
                                  from = "2022-01-01",
                                  apikey = apikey)

```

## Development
If you encounter problems or if you have suggestions for improvement, let me know or, ideally, open an issue. Also, I am looking forward to pull requests. 

## Disclaimer
This package has been developed independently of and is not in any way associated to GIE or the AGSI/ALSI+ transparency platform. Some parts of or calls to the API may exhibit unforeseen behaviour due to the API still being subject to changes. 
